Gromov--Witten theory of hypersurfaces in genus one

### 2019-03-01 GenusOneRelativeMaps.tex
started step 4 - rubber recursion

\subsection*{Step 4} Finally, suppose we are given a special reduced rubber invariant to compute. So we have a reduced rubber space $\VZ_{1,\alpha}(P|Y_0+Y_\infty,\beta)_{\Gm}$ together with an insertion $\gamma$ satisfying the conditions outlined above. We first note that if there exists a relative marking with no insertion, we may forget this marking and identify the rubber invariant with a recursively-known absolute invariant (up to some boundary terms which can also be computed recursively). This approach will suffice for the majority of computations.

[TO BE FINISHED]

>>>>

\subsection*{Step 4} Finally, we show how to compute any rubber invariant (even if, for the recursion to keep going, we would only need the special ones). From the previous steps we effectively use only the knowledge of absolute invariants of $Y$ with not higher $(d,n)$ than the current step. We proceed by induction on $(d,n,n^\prime)$, where $n^\prime$ is the number of markings with non-zero tangency requirement.

Recall that there is a map $\epsilon\colon \VZ_{1,\alpha}(P|H_b+H_t)^\sim\to \VZ_{1,\alpha}(H)$ projecting down to the base and stabilising the curve, i.e. collapsing rational bridges; we always assume our insertions to be pulled back along $\epsilon$, in particular the psi classes are \emph{collapsed}. Forgetting a marking may be thought of as a map to a different rubber space only if it is a non-relative one; otherwise it will always be considered as a map to the collapsed space.

We start by recalling some useful constructions on the space of maps to rubber \cite{EKatz}.
\begin{enumerate}[leftmargin=15pt]
 \item The moduli space of targets is isomorphic to the semistable locus in $\mathfrak M_{0,2}$. Pulling back the cotangent line bundles at $p_b,\ p_t$ we obtain $\mathbb L_b,\ \mathbb L_t$, whose $c_1$ are the so-called \emph{target psi classes} $\Psi_b,\ \Psi_t$. The subscripts stand for \emph{bottom} and \emph{top}.
 \item The line bundle $\mathrm{Split}=\mathbb L_b\otimes\mathbb L_t$ admits a section, whose vanishing locus consists precisely of those maps whose target has more than one irreducible component. This can be seen as the pullback of the topological recursion relation on the moduli space of targets. $\mathrm{Split}$ admits a tropical description as the piecewise-linear function assigning the sum of the finite edge-lengths of the target.
 \item Whenever there is a marking $x_k$ of positive tangency $\alpha_k$ (i.e. mapping to $D_t$), we may express the target psi class $\Psi_t$ as $\alpha_k\hat\psi_k+\ev_k^*H$. Importantly, here $\hat\psi_k$ denotes the expanded psi class at $x_k$, which differs from the collapsed psi class - that we usually consider - on the locus $\delta_{k,\rm{triv}}$ where $x_k$ lies on a trivial bubble. Similarly $\Psi_b=-\alpha_h\hat\psi_h-\ev_h^*H$ for a marking $x_h$ with $\alpha_h<0$. See also \cite[Construction 5.1.17]{GathmannThesis}.
 \item\label{claim:trivialPsib} $\Psi_b=0$ whenever there are no negative markings.
 \item For every non-relative marking (with $\alpha_i=0$), there is a line bundle $L_{i,\rm{not\ bot}}=\mathbb L_b\otimes\ev_i^*\OO(H)$ with a section vanishing on the locus where (the target splits and) $x_i$ does not map to the bottom component of the target. This line bundle has a tropical interpretation by taking the distance between $\phi(v_i)$ and the left-most component in the target, where $v_i$ is the vertex of the curve supporting the $i$-th flag.

 \noindent Similarly, $L_{i,\rm{not\ top}}=\mathbb L_t\otimes\ev_i^*\OO(-H)$ has a section cutting out the locus where $x_i$ does not map to the top component of the target.
\end{enumerate}

\emph{First reduction}: assume there is an index (wlog $1$) such that no insertion involves $x_1$. Then we would like to replace the whole insertion $\gamma$ by $\fgt_{x_1}^*(\gamma)$. We only need to notice that, for $k\neq1$, the difference $\psi_k-\fgt_{x_1}^*(\psi_k)$ is supported on the locus where $x_1$ and $x_k$ alone lie on a rational component of trivial horizontal degree; the invariants arising from $\delta_{1,k}$ can be computed recursively as they have lower $n$.

The base for the induction is the case of a single relative marking $x_1$ with full tangency $\alpha_1=d$. Write the insertion as $\gamma=\tau_{b_1}(\gamma_1)\tau(\gamma)_{\geq 2}$. We shall add a non-relative marking $x_0$ and, starting from the equality $c_1(L_{0,\rm{not\ bot}})\cdot[\VZ_{1,\alpha}(P|H_0+H_\infty)^\sim]=\ldots$ (the boundary classes described above), intersect it with $\bar\gamma=\fgt_{x_1}^*\left(\psi_0^{b_1}\ev_0^*(\gamma_1)\tau(\gamma)_{\geq 2}\right)$. \begin{comment}{$\bar\gamma=\fgt_{x_1}^*\left(\psi_0^{b_1}\ev_0^*(\gamma_1)\tau(\gamma)_{\geq 2}\right)$\end{comment}
On the right-hand side, all the not recursively known terms consist of an elliptic curve of degree $d$ in the bottom level, and a rational tail of horizontal degree $0$ in the top level. The rational tail supports $x_0$ by assumption, and it must support $x_1$ as well by balancing: if any other marking is on it, the invariant is recursively known (lower $n$); otherwise, we recover the invariant we are trying to compute (up to corrections from comparing $\psi_k$ and $\fgt_{x_1}^*\psi_k$, which again involve invariants with lower $n$). On the left hand side, write $c_1(L_{0,\rm{not\ bot}})=\Psi_b-\ev_0^*H=-\ev_0^*H$ by \ref{claim:trivialPsib} above. We can therefore compute the left-hand side by forgetting $x_1$ and the \emph{first reduction}: the result is the absolute invariant $-d^2\langle\tau_{b_1}(H\cdot\gamma_1)\tau(\gamma)_{\geq2}\rangle^{H,\rm{red}}_{1,n,d}$, which is recursively known.
\newpage
